The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-01 paper.tex
new figure

\section{What does Anthropomorphism mean?}
\label{sec:anthropomorphism}

@@

\hl{Julia: I will make a draft of a figure here}

@@

The underlying central point about the human-like design of robots lies in the 
implications that this design holds, and the suggestions for interactions that
it makes to the human user. 
In other words, the appearance 
and function of a product, whether it is an appliance or a humanoid robot, affects how 
people perceive it, interact with it, and build long-term relationships with 
it \citep{bartneck_shaping_2004}. 
The principle of human-like form is probably most prominent in the design of robots, 
due to several reasons, such as to foster efficient interaction and increase 
acceptance. An anthropomorphic design of a robot encourages people to reason about this 
unfamiliar system \emph{as if} it was a human. To increase this tendency, robots are 
designed with anthropomorphic (or lifelike) forms, or are able to display and use 
human-social cues when interacting with humans.
However, anthropomorphic robot designs have been controversially discussed and received 
remarkable attention since the early 1970ies, related to \cite{mori_uncanny_1970} 
theory of the \textit{Uncanny Valley}, see Figure \ref{fig:uncanny_valley}.

\begin{figure}\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.75]{uncanny-valley.jpg}
% figure caption is below the figure
 \caption{Mashiro Mori, illustration of the \textit{Uncanny Valley} hypothesis. The curve models the user's perceived affinity with the artifact, as its human likeness is increased. Up to a certain (not yet defined) point, a person perceives more affinity in a more human-like agent; however with an abrupt shift towards a negative perception as soon as the agent surpasses this point, of resembling a human too closely but not fully. Originally from 1970; picture from a manuscript version of \cite{mori_uncanny_2012}}
 \label{fig:uncanny_valley}       % Give a unique label
 \end{figure}

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
Apart from this, it is suggested that an increased human-likeness of a robot (up to a 
certain threshold) will generally lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \citep{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
that humans tend to project everyday 
expectations of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \citep{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus, the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot. 
These experiences mark the transition between an immediate \textit{in-the-moment} and more thoughtful \textit{reflective} \citep{takayama_perspectives_2012} perception of, and interaction with the robot. As such, a more human-centered perspective needs to be applied, and we will come back to this in Section \ref{sec:human-centered}.

>>>>

\section{Why do humans anthropomorphize?}
\label{sec:anthropomorphism}

@@

\begin{figure}[b]
    \centering
    \includegraphics[width=0.75\columnwidth]{anthropo.pdf}
    \caption{what is anthropomorphism? \hl{we need a cool caption for this that includes terms such as factor, element, influence, etc. :-) }}

    \label{fig:anthropofig}
\end{figure}

@@

The underlying central point about the human-like design of robots lies in the 
implications that this design holds, and the suggestions for interactions that
it makes to the human user. 
In other words, the appearance 
and function of a product, whether it is an appliance or a humanoid robot, affects how 
people perceive it, interact with it, and build long-term relationships with 
it \citep{bartneck_shaping_2004}. 
The principle of human-like form is probably most prominent in the design of robots, 
due to several reasons, such as to foster efficient interaction and increase 
acceptance. An anthropomorphic design of a robot encourages people to reason about this 
unfamiliar system \emph{as if} it was a human. To increase this tendency, robots are 
designed with anthropomorphic (or lifelike) forms, or are able to display and use 
human-social cues when interacting with humans.
However, anthropomorphic robot designs have been controversially discussed and received 
remarkable attention since the early 1970ies, related to Mori's
theory of the \textit{Uncanny Valley} \citep{mori_uncanny_1970}.
%see Figure \ref{fig:uncanny_valley}.

%\begin{figure}\centering
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[scale=0.75]{uncanny-valley.jpg}
%% figure caption is below the figure
% \caption{Mashiro Mori, illustration of the \textit{Uncanny Valley} hypothesis. The curve models the user's perceived affinity with the artifact, as its human likeness is increased. Up to a certain (not yet defined) point, a person perceives more affinity in a more human-like agent; however with an abrupt shift towards a negative perception as soon as the agent surpasses this point, of resembling a human too closely but not fully. Originally from 1970; picture from a manuscript version of \cite{mori_uncanny_2012}}
% \label{fig:uncanny_valley}       % Give a unique label
% \end{figure}

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
Apart from this, it is suggested that an increased human-likeness of a robot (up to a 
certain threshold) will generally lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970}.
% (see Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \citep{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
that humans tend to project everyday 
expectations of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \citep{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus, the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot. 
These experiences mark the transition between an immediate \textit{in-the-moment} and more thoughtful \textit{reflective} \citep{takayama_perspectives_2012} perception of, and interaction with the robot. As such, a more human-centered perspective needs to be applied, and we will come back to this in Section \ref{sec:human-centered}.
